Hydra Battles and Cie

### 2021-03-12 doc/part-hydras.tex
consistent use of capitalization in headers

\subsection*{State of the Development}
The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.

@@

\section*{Future Work (projects)}
\index{hydras}{Projects}

@@

\subsection*{Main References}

@@

\subsection{The Rules of the Game}

@@

\section{Hydras and their Representation in \emph{Coq}}
\label{sec:orgheadline48}

@@

\subsubsection{Recursive Functions on type \texttt{Hydra}}
\label{sec:orgheadline41}
\label{sec:hsize-def}

@@

\subsection{Induction Principles for Hydras}
\label{sec:orgheadline42}

@@

\subsubsection{A failed Attempt}

@@

\subsubsection{A Principle of Mutual Induction}
In order to get an appropriate induction scheme for the types 
\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.

@@

\subsubsection{A Correct Proof}

@@

\section{Relational Description of Hydra Battles}

@@

\subsection{Chopping off a Head at Distance 1 from the Foot (Relation  R1)}

@@

\subsection{Chopping off a Head at Distance \texorpdfstring{$\geq 2$}{>= 2} from the Foot (relation R2) }

@@

\subsection{Relation Associated with a Round}

@@

\subsection{Rounds and Battles}

@@

\subsection{Classes of Battles}
\label{sect:battle-classes}

@@

\subsection{Big Steps}

@@

\section{A Long Battle}
\label{sect:big-battle}

@@

\subsection{The Beginning of Hostilities}
During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.

@@

\subsection{Looking for Regularities}

@@

\subsection{Giant Steps}

@@

\subsection{A Minoration Lemma}

@@

\section{Generic Properties}

@@

\subsection{A  Small Proof of Impossibility}
\index{coq}{Proofs of impossibility}

@@

\section{The Mathematical Point of View}

\subsection{Well-ordered Sets}
Let us start with some definitions.
A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
\begin{description}
\item[irreflexivity] : $\forall x\in A, x\not< x$
\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$. In contrast, \coq's standard library contains
an inductive definition of a predicate \texttt{Acc} which allows us to write 
constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
\end{description}

@@

\subsection{Ordinal Numbers}

@@

\section{Ordinal Numbers in Coq}

@@

% \section{Countable Ordinals}

@@

\subsection{A Class for Ordinal Notations}

@@

\subsection{Ordinal Notations and Measures for Proving Termination}
\label{sect:measure-ON}

@@

\section{Example: the Ordinal \texorpdfstring{$\omega$}{omega}}

@@

\section{Sum of  two Ordinal Notations}

@@

\subsection{The Ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}

@@

\section{Limits and Successors}

@@

\subsection{Limits and Successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}

@@

\section{Product of Ordinal Notations}

@@

\section{The Ordinal \texorpdfstring{$\omega^2$}{omega2}}

@@

\subsection{A Proof of Termination using \texorpdfstring{$\omega^2$}{omega^2}} 
\label{sect:merge-example}.

@@

\subsection{Yet Another  Proof of Impossibility}
\label{omega2-case}

@@

\subsubsection{Proof of the Inequality \texttt{m small\_h o< m big\_h}}

@@

\subsubsection{Proof of the Inequality \texttt{m big\_h o<= m small\_h} }

@@

\subsubsection{End of the Proof}
From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).

@@

\section{A Notation for Finite Ordinals}

@@

\section{Comparing two Ordinal Notations}

@@

\section{Comparing an Ordinal Notation with Schütte's Model}

Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model). This would be a relative proof of correctenss of the considered  ordinal  notation.

@@

\section{Isomorphism of Ordinal Notations}

@@

\section{Other Ordinal Notations}

@@

\section{The Ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
\label{sec:epsilon0-intro}

\subsection{Cantor Normal Form}
\index{maths}{Cantor normal form}

@@

\subsection{A Data Type for  Ordinals in Cantor Normal Form}
\label{sec:orgheadline72}
\label{sec:T1-inductive-def}

@@

\subsubsection{Finite Ordinals}
\label{sec:orgheadline67}

@@

\subsubsection{The Ordinal \(\omega\)}
\label{sec:orgheadline68}

@@

\subsubsection{The Ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
\label{sect:notation-phi0}
We provide also a notation for ordinals of the form $\omega^\alpha$.

@@

\subsubsection{The Hierarchy of \(\omega\)-towers:}
\label{sec:orgheadline71}

@@

\subsection{Pretty Printing Ordinals in Cantor Normal Form}
\label{sect:ppT1}
\index{hydras}{Library Epsilon0!Types!ppT1}

@@

\subsection{Comparison between Ordinal Terms}
\label{sec:orgheadline73}

@@

\subsection{Making Normality Implicit}
  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
A simple way to do this is to consider statements of the form 
\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.

@@

\subsubsection{A Sigma Type for \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\subsection{Syntactic Definition of Limit and Successor Ordinals}

@@

\subsubsection{Addition and Multiplication}

@@

\section{Well-foundedness and Transfinite Induction}

@@

\subsection{About  Well-foundedness}
\label{sec:orgheadline82}
   In order to use \texttt{T1} for proving termination results,
we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.

@@

\subsubsection{A First Attempt}
\label{sec:orgheadline77}
\index{coq}{Well-founded induction}

@@

\subsubsection{Using a Stronger Inductive Predicate.}
\label{sec:orgheadline78}
  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

@@

\subsection{An Ordinal Notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\section{A Variant for Hydra Battles}

@@

\subsection{Natural Sum (a.k.a. Hessenberg's  Sum)}
\label{sec:orgheadline87}
\label{hydra-variant}

@@

\subsubsection{Rewriting Lemmas}
\label{sec:orgheadline86}

@@

\index{hydras}{Projects}
\begin{project}
Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
(\texttt{coq-equations}, \texttt{Function}, etc.).
\end{project}
\subsection{More Theorems on Hessenberg's Sum}

@@

\subsubsection{Commutativity, Associativity}

@@

\subsection{A Measure for Hydra-battle Termination}

@@

\subsubsection{Mathematical Definition of Canonical Sequences}

@@

\subsubsection{Canonical Sequences in Coq}
\index{hydras}{Library Epsilon0!Functions!canon}
\index{hydras}{Library Epsilon0!Functions!canonS}

@@

\subsection{Basic Properties of Canonical Sequences}

@@

\subsubsection{Canonical Sequences and the Order $<$}

@@

\subsubsection{Limit Ordinals are truly Limits}
The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.

@@

\subsection{Formal Definition}

@@

\subsection{Existence of a Path}

@@

\subsection{Paths and Hydra Battles}
\label{KS-o2h}

@@

\section{A  Proof of Impossibility}

@@

\subsection{The case of Free Battles}
\label{sec:free-battles-case}
Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles). The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.

@@

\section{The Case of Standard Battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

\subsection{Paths with a Constant Index}

@@

\subsubsection{Sketch of Proof of \texttt{Cor12}}
\index{maths}{Transfinite induction}

@@

\subsection{Casting Paths with a Constant Index into a Standard Path}

@@

\subsubsection{Sketch of Proof of \texttt{constant\_to\_standard\_path}}

@@

\subsubsection*{Préliminaries}

@@

\subsection{Back to Hydras}
\label{sec:standard-battles-cases}
We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
be consulted in the module 
\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
% ICI !
Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
{\texttt{Epsilon0.Paths}}   and 
\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
{\texttt{constant\_to\_standard\_path}},
we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.

@@

\section{Length of Minimal Large Sequences}

@@

\subsection{Formal Specification}

@@

\subsection{Abstract Properties}

@@

\subsubsection{The  Ordinal Zero}
\label{sect:L-spec-zero}
The base case is directly a consequence of the specification.

@@

\subsubsection{Successor Ordinals}
\label{sect:L-spec-succ}
Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
$\beta$ at $k+2$ to $0$. 
By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.

@@

\subsubsection{Limit Ordinals}
\label{sect:L-spec-lim}

@@

\subsection{First Results}

@@

\subsubsection{Finite Ordinals}

@@

\subsubsection{The First Limit Ordinal  \texorpdfstring{$\omega$}{omega}}

@@

\subsubsection{Going Further}
Let us consider a last example, ``computing'' $L_{\omega^3}$.
Since the canonical sequence associated with this ordinal is composed of the
$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.

@@

\subsubsection{Back to Hydra Battles}

@@

\section{The Wainer-Hardy Hierarchy (Functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}

@@

\subsection{Hardy Functions in \texttt{Coq}}

@@

\subsection{First  Steps of the Hardy hierarchy}
Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.

\subsubsection{Finite Ordinals}

@@

\subsubsection{New Limits}

@@

\subsubsection{A Numerical Example}

@@

\subsection{Abstract Properties of Hardy functions}
~\label{sect:H-alpha-prop}

@@

\subsubsection{Back to Hydras}

@@

\section{The Wainer Hierarchy (Functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
\label{sect:wainer}

@@

\section{Declarations and Axioms}

@@

\section{Additional  Axioms}

@@

\subsubsection{Classical Logic}

@@

\subsubsection{Description Operators}

@@

\subsubsection{The Definition of zero}

@@

\section{The  Successor Function}

@@

\section{Finite Ordinals}

@@

\section{The Definition of \texttt{omega}}
In order to define $\omega$, the first infinite ordinal, we use an operator which
``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
For that purpose, we first use a predicate:
(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
upper bound  of $X$ (with respect to \textit{lt}).

@@

\subsection{Ordering Functions and Ordinal Addition}

@@

\subsection{Ordinal Addition}

@@

\subsubsection{Multiplication by a Natural Number}

@@

\section{The Exponential of Basis \texorpdfstring{$\omega$}{omega}}

@@

\subsection{Additive Principal Ordinals}

@@

\subsection{The Function \texttt{phi0}}

@@

\subsection{Omega-towers and the Ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\subsection{Properties of the Set  \texttt{AP}}

@@

\subsubsection{Properties of the Function \texorpdfstring{$\phi_0$}{phi0}}

@@

\section{Critical Ordinals}

@@

\subsection{A flavor of Infinity}

@@

\section{Cantor Normal Form}

@@

\section{An Embedding of \texttt{T1} into \texttt{Ord}}

@@

\section{Related Work}

@@

\section{The Type \texttt{T2} of Ordinal Terms}

@@

\section{How Big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}

@@

\section{Veblen Normal Forms}
\begin{definition}
  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
 \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
\end{definition}

@@

\subsection{Length of a Term}

@@

\section{Main Functions on \texttt{T2}}

@@

\subsection{The Veblen Function \texorpdfstring{$\phi$}{\texttt{phi}}}

@@

\section{An Ordinal Notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}

>>>>

\subsection*{State of the development}
The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on  notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.

@@

\section*{Future work (projects)}
\index{hydras}{Projects}

@@

\subsection*{Main references}

@@

\subsection{The rules of the game}

@@

\section{Hydras and their representation in \emph{Coq}}
\label{sec:orgheadline48}

@@

\subsubsection{Recursive functions on type \texttt{Hydra}}
\label{sec:orgheadline41}
\label{sec:hsize-def}

@@

\subsection{Induction principles for hydras}
\label{sec:orgheadline42}

@@

\subsubsection{A failed attempt}

@@

\subsubsection{A Principle of mutual induction}
In order to get an appropriate induction scheme for the types 
\texttt{Hydra} and \texttt{Hydrae}, we can use  \coq{}'s  command \texttt{Scheme}.

@@

\subsubsection{A Correct proof}

@@

\section{Relational description of hydra battles}

@@

\subsection{Chopping off a head at distance 1 from the foot (relation  R1)}

@@

\subsection{Chopping off a head at distance \texorpdfstring{$\geq 2$}{>= 2} from the foot (relation R2) }

@@

\subsection{Binary relation associated with a round}

@@

\subsection{Rounds and battles}

@@

\subsection{Classes of battles}
\label{sect:battle-classes}

@@

\subsection{Big steps}

@@

\section{A Long battle}
\label{sect:big-battle}

@@

\subsection{The beginning of hostilities}
During the two first rounds, our hydra loses its two rightmost heads.  Figure~\vref{fig:hinit-plus2} shows the state of the hydra   just before the third round.

@@

\subsection{Looking for regularities}

@@

\subsection{Giant steps}

@@

\subsection{A minoration lemma}

@@

\section{Generic properties}

@@

\subsection{A  small proof of impossibility}
\index{coq}{Proofs of impossibility}

@@

\section{The mathematical point of view}

\subsection{Well-ordered sets}
Let us start with some definitions.
A  \emph{well-ordered set} is a set provided with a binary relation $<$ which has the following properties.
\begin{description}
\item[irreflexivity] : $\forall x\in A, x\not< x$
\item[transitivity] : $\forall x\,y\,z\in A, x<y \Rightarrow y<z \Rightarrow x<z$
\item[trichotomy]: $\forall x\,y\in A, x<y \vee x = y \vee y < x$
\item[well foundedness]: $<$ is well-founded (every element of $A$ is accessible)\footnote{In classical mathematics, we would say that there is no infinite sequence $a_1>a_2> \dots a_n> a_{n+1}\dots$ in $A$. In contrast, \coq's standard library contains
an inductive definition of a predicate \texttt{Acc} which allows us to write 
constructive proofs of accessibility (See \href{https://coq.inria.fr/distrib/current/stdlib/Coq.Init.Wf.html}{Coq.Init.Wf}).}.
\end{description}

@@

\subsection{Ordinal numbers}

@@

\section{Ordinal numbers in Coq}

@@

% \section{Countable ordinals}

@@

\subsection{A class for ordinal notations}

@@

\subsection{Ordinal notations and  termination measures}
\label{sect:measure-ON}

@@

\section{Example: the ordinal \texorpdfstring{$\omega$}{omega}}

@@

\section{Sum of  two ordinal notations}

@@

\subsection{The ordinal \texorpdfstring{$\omega+\omega$}{omega + omega}}

@@

\section{Limits and successors}

@@

\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}

@@

\section{Product of ordinal notations}

@@

\section{The ordinal \texorpdfstring{$\omega^2$}{omega2}}

@@

\subsection{A proof of termination using \texorpdfstring{$\omega^2$}{omega^2}} 
\label{sect:merge-example}.

@@

\subsection{Yet another  proof of impossibility}
\label{omega2-case}

@@

\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}

@@

\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }

@@

\subsubsection{End of the proof}
From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).

@@

\section{A notation for finite ordinals}

@@

\section{Comparing two ordinal notations}

@@

\section{Comparing an ordinal notation with Schütte's model}

Finally, it may be interesting to compare an ordinal notation with the more theoretical model from Schütte (well, at least with our formalization of that model). This would be a relative proof of correctness of the considered  ordinal  notation.

@@

\section{Isomorphism of ordinal notations}

@@

\section{Other ordinal notations}

@@

\section{The ordinal \texorpdfstring{\(\epsilon_0\)}{epsilon0}}
\label{sec:epsilon0-intro}

\subsection{Cantor normal form}
\index{maths}{Cantor normal form}

@@

\subsection{A data type for  ordinals in Cantor normal form}
\label{sec:orgheadline72}
\label{sec:T1-inductive-def}

@@

\subsubsection{Finite ordinals}
\label{sec:orgheadline67}

@@

\subsubsection{The ordinal \(\omega\)}
\label{sec:orgheadline68}

@@

\subsubsection{The ordinal \(\omega^\alpha\), a.k.a. \(\phi_0(\alpha)\)}
\label{sect:notation-phi0}
We provide also a notation for ordinals of the form $\omega^\alpha$.

@@

\subsubsection{The hierarchy of \(\omega\)-towers:}
\label{sec:orgheadline71}

@@

\subsection{Pretty-printing ordinals in Cantor normal form}
\label{sect:ppT1}
\index{hydras}{Library Epsilon0!Types!ppT1}

@@

\subsection{Comparison between ordinal terms}
\label{sec:orgheadline73}

@@

\subsection{Making normality implicit}
  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
A simple way to do this is to consider statements of the form 
\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.

@@

\subsubsection{A sigma-type for \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\subsection{Syntactic definition of limit and successor ordinals}

@@

\subsubsection{Addition and multiplication}

@@

\section{Well-foundedness and transfinite induction}

@@

\subsection{About  well-foundedness}
\label{sec:orgheadline82}
   In order to use \texttt{T1} for proving termination results,
we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.

@@

\subsubsection{A first attempt}
\label{sec:orgheadline77}
\index{coq}{Well-founded induction}

@@

\subsubsection{Using a stronger inductive predicate.}
\label{sec:orgheadline78}
  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

@@

\subsection{An ordinal notation for  \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\section{A Variant for hydra battles}

@@

\subsection{Natural sum (a.k.a. Hessenberg's  sum)}
\label{sec:orgheadline87}
\label{hydra-variant}

@@

\subsubsection{Rewriting lemmas}
\label{sec:orgheadline86}

@@

\index{hydras}{Projects}
\begin{project}
Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
(\texttt{coq-equations}, \texttt{Function}, etc.).
\end{project}
\subsection{More theorems on Hessenberg's sum}

@@

\subsubsection{Commutativity, associativity}

@@

\subsection{A termination measure for hydra battles }

@@

\subsubsection{Mathematical definition of canonical sequences}

@@

\subsubsection{Canonical sequences in Coq}
\index{hydras}{Library Epsilon0!Functions!canon}
\index{hydras}{Library Epsilon0!Functions!canonS}

@@

\subsection{Basic properties of canonical sequences}

@@

\subsubsection{Canonical sequences and the order $<$}

@@

\subsubsection{Limit ordinals are truly limits}
The following theorem states that any limit ordinal $\lambda<\epsilon_0$ 
is the limit of the sequence \showmath{\canonseq{\lambda}{i}\;(1\le i)}.

@@

\subsection{Formal definition}

@@

\subsection{Existence of a path}

@@

\subsection{Paths and hydra battles}
\label{KS-o2h}

@@

\section{A  proof of impossibility}

@@

\subsection{The case of free battles}
\label{sec:free-battles-case}
Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles). The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.

@@

\section{The case of standard battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

\subsection{Paths with a constant index}

@@

\subsubsection{Sketch of proof of \texttt{Cor12}}
\index{maths}{Transfinite induction}

@@

\subsection{Casting paths with a constant index into a standard path}

@@

\subsubsection{Sketch of proof of \texttt{constant\_to\_standard\_path}}

@@

\subsubsection*{Preliminaries}

@@

\subsection{Back to hydras}
\label{sec:standard-battles-cases}
We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
be consulted in the module 
\url{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}.
Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
% ICI !
Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
{\texttt{Epsilon0.Paths}}   and 
\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
{\texttt{constant\_to\_standard\_path}},
we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.

@@

\section{Length of minimal large sequences}

@@

\subsection{Formal specification}

@@

\subsection{Abstract properties}

@@

\subsubsection{The  ordinal zero}
\label{sect:L-spec-zero}
The base case is directly a consequence of the specification.

@@

\subsubsection{Successor ordinals}
\label{sect:L-spec-succ}
Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
$\beta$ at $k+2$ to $0$. 
By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.

@@

\subsubsection{Limit ordinals}
\label{sect:L-spec-lim}

@@

\subsection{First results}

@@

\subsubsection{Finite ordinals}

@@

\subsubsection{The first limit ordinal  \texorpdfstring{$\omega$}{omega}}

@@

\subsubsection{Going further}
Let us consider a last example, ``computing'' $L_{\omega^3}$.
Since the canonical sequence associated with this ordinal is composed of the
$\omega^2\times i\;(i\in\mathbb{N}_1)$, we have to study this sequence.

@@

\subsubsection{Back to hydra battles}

@@

\section{The Wainer-Hardy hierarchy (functions \texorpdfstring{$H_\alpha$}{\texttt{H\_alpha}})}

@@

\subsection{Hardy functions in \texttt{Coq}}

@@

\subsection{First  steps of the Hardy hierarchy}
Using rewrite rules from \texttt{H\_eq1} to \texttt{H\_eq4}, we can explore the functions $H_\alpha$ for some small values of $\alpha$.

\subsubsection{Finite ordinals}

@@

\subsubsection{New limits}

@@

\subsubsection{A numerical example}

@@

\subsection{Abstract properties of Hardy functions}
~\label{sect:H-alpha-prop}

@@

\subsubsection{Back to hydras}

@@

\section{The wainer hierarchy (functions \texorpdfstring{$F_\alpha$}{F\_alpha})}
\label{sect:wainer}

@@

\section{Declarations and axioms}

@@

\section{Additional  axioms}

@@

\subsubsection{Classical logic}

@@

\subsubsection{Description operators}

@@

\subsubsection{The definition of zero}

@@

\section{The  successor function}

@@

\section{Finite ordinals}

@@

\section{The definition of \texttt{omega}}
In order to define $\omega$, the first infinite ordinal, we use an operator which
``returns'' the least upper bound (if it exists) of a subset $X\subseteq \mathbb{O}$.
For that purpose, we first use a predicate:
(\texttt{is\_lub $D$ \textit{lt} $X$ $a$}) if $a$ belongs to $D$ and is the least 
upper bound  of $X$ (with respect to \textit{lt}).

@@

\subsection{Ordering functions and ordinal addition}

@@

\subsection{Ordinal addition}

@@

\subsubsection{Multiplication by a natural number}

@@

\section{The exponential of basis \texorpdfstring{$\omega$}{omega}}

@@

\subsection{Additive principal ordinals}

@@

\subsection{The function \texttt{phi0}}

@@

\subsection{Omega-towers and the ordinal \texorpdfstring{$\epsilon_0$}{epsilon0}}

@@

\subsection{Properties of the set  \texttt{AP}}

@@

\subsubsection{Properties of the function \texorpdfstring{$\phi_0$}{phi0}}

@@

\section{Critical ordinals}

@@

\subsection{A flavor of infinity}

@@

\section{Cantor normal form}

@@

\section{An embedding of \texttt{T1} into \texttt{Ord}}

@@

\section{Related work}

@@

\section{The type \texttt{T2} of ordinal terms}

@@

\section{How big is \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}?}

@@

\section{Veblen normal form}
\begin{definition}
  A term of the form $\psi(\alpha_1,\beta_1)\times n_1+ \psi(\alpha_2,\beta_2)\times n_2+\dots+\psi(\alpha_k,\beta_k)\times n_k$ is said to be in
 \emph{[Veblen] normal form} if for every $i<n$, $\psi(\alpha_i,\beta_i)<\psi(\alpha_{i+1},\beta_{i+1})$, all the $\alpha_i$ and $\beta_i$ are in normal form, and all the $n_i$ are strictly positive integers.
\end{definition}

@@

\subsection{Length of a term}

@@

\section{Main functions on \texttt{T2}}

@@

\subsection{The Veblen function \texorpdfstring{$\phi$}{\texttt{phi}}}

@@

\section{An ordinal notation for \texorpdfstring{$\Gamma_0$}{\texttt{Gamma0}}}
